---
title: "Issues with 64-bit integers in R"
author: "MJ Schuemie"
date: "1/29/2021"
output:
  pdf_document: default
---

```{r warn=FALSE, message=FALSE}
library(dplyr)
```

# Dangerous

Things that fail without immediate warning

## For in loop

My favorite loop pattern leads to bogus numbers:

```{r}
for (i in bit64::as.integer64(1:3))
  print(i)
```

## Any apply

sapply, lapply, mapply, etc. lead to bogus numbers:

```{r}
sapply(bit64::as.integer64(1:3), print)
```

Should use purrr::map instead:
```{r}
purrr::map(bit64::as.integer64(1:3), print)
```

## The %in% operator when other variable is not integer64

```{r}
as.numeric(1) %in% bit64::as.integer64(1:3)
as.integer(1) %in% bit64::as.integer64(1:3)
bit64::as.integer64(1) %in% as.numeric(1:3)
bit64::as.integer64(1) %in% as.integer(1:3)
```

## RJSONSIO::toJson

Converts to bogus numbers:

```{r}
RJSONIO::toJSON(list(i = bit64::as.integer64(1)))
```


# Annoying

Things that throw an error, but can be avoided by explicit casting.

## Must explictly cast when reading from CSV

```{r}
file <- tempfile()
df <- data.frame(i = bit64::as.integer64(1))
readr::write_csv(df, file)

df2 <- readr::read_csv(file, col_types = c(i = "c")) %>%
    mutate(i = bit64::as.integer64(.data$i))
df2
```

## Bind_rows cannot combine integer64 with numeric

```{r eval=FALSE}
# This throws and error:
bind_rows(data.frame(i = bit64::as.integer64(1)),
          data.frame(i = as.numeric(2)))
```

## RSQLite (and Andromeda) only knows 64-bit integers

For this reason, the type of integer that is returned depends on the value:

```{r}
df <- data.frame(key = c(1,2),
                 value = bit64::as.integer64(c(2^2, 2^62)))
andr <- Andromeda::andromeda(table = df)
andr$table %>%
  filter(key == 1)

andr$table %>%
  filter(key == 2)
```
